---
title: "PCS Documentation"
author: ""
date: "`r format(Sys.time(), '%B %d, %Y')`"
header-includes:
    - \usepackage{float}
    - \usepackage{amsmath}
    - \usepackage{gensymb}
    - \usepackage{tcolorbox}
output:
  rmdformats::material:
    fig_caption: true
    code_folding: hide
    number_sections: true
    use_bookdown: true
    fig_width: 10
    fig_height: 8
    lightbox: true
    code_download: true
    includes:
      before_body: html/setup.html
  pdf_document:
    number_sections: true
params:
  X_filepath:
    label: "X Data"
    # value: NULL
    value: data/tcga_brca_array_data.rds
    input: file 
  y_filepath:
    label: "y data"
    # value: NULL
    value: data/tcga_brca_subtypes.rds
    input: file
  train_prop:
    label: "Training data proportion"
    value: 0.6
    input: numeric
  valid_prop:
    label: "Validation data proportion"
    value: 0.2
    input: numeric
  test_prop:
    label: "Test data proportion"
    value: 0.2
    input: numeric
  modeling_pkg:
    label: "Modeling Package"
    value: "caret"
    input: select
    choices: ["caret", "h2o", "tidymodels"]
  seed:
    label: "Random Seed"
    value: 12345
    input: numeric
css: [css/custom_material_rmd_theme.css, css/custom_checkbox.css]
# runtime: shiny
---

```{r setup, echo = FALSE, message=FALSE, warning=FALSE}
source("scripts/rmd-setup.R", local = knitr::knit_global())

# load in packages
library(tidyverse)
library(skimr)
library(R.utils)
library(datasets)
library(ggplot2)
library(dplyr)
library(tidyr)
library(reshape2)
for (f in list.files("R", pattern = ".R", full.names = TRUE)) {
  source(f, chdir = T)
}

# set seed
set.seed(params$seed)

# validate inputs
validateDataSplit(params$train_prop, params$valid_prop, params$test_prop)
if (params$modeling_pkg != "caret") {
  stop("Only caret modeling backend has been implemented.")
}

# load data in
X <- uploadFile(params$X_filepath)
y <- uploadFile(params$y_filepath)
#TODO: validate data better
validateData(X, y)

# initialize counter for subchunkify
chunk_idx <- 1
```

# Domain problem formulation

What is the real-world question? This could be hypothesis-driven or discovery-based.

```{asis, help = TRUE}
Some advice
```

```{asis, interactive_text = TRUE}

```

Why is this question interesting and important? What are the implications of better understanding this data?

```{asis, interactive_text = TRUE}

```

Briefly describe any background information necessary to understand this problem.

```{asis, help = TRUE}
Some advice
```

```{asis, interactive_text = TRUE}

```

Briefly describe how this question can be answered in the context of a model or analysis.

```{asis, interactive_text = TRUE}

```

Outline the rest of the report/analysis.

```{asis, interactive_text = TRUE}

```

# Data

What is the data under investigation? Provide a brief overview/description of the data.

```{asis, interactive_text = TRUE}

```

Describe how your data connects to the domain problem.

```{asis, interactive_text = TRUE}

```

## Data Collection

How was the data collected or generated (including details on the experimental design)? Be as transparent as possible so that conclusions made from this data are not misinterpreted down the road.

```{asis, interactive_text = TRUE}

```

Describe any limitations when using the data to answer the domain problem of interest.

```{asis, interactive_text = TRUE}

```

Where is the data stored, and how can it be accessed by others (if applicable)?

```{asis, interactive_text = TRUE}

```

## Data Splitting

TODO: add advice for possible data splits, AK getting nice figure together

Decide on the proportion of data in each split.

Decide on the "how" to split the data (e.g., random sampling, stratified sampling, etc.), and explain why this is a reasonable way to split the data.

```{asis, interactive_text = TRUE}

```

Split the data into a training, validation, and test set.

```{r split-data}
data_split <- basicDataSplit(X = X, y = y, 
                             train_prop = params$train_prop, 
                             valid_prop = params$valid_prop, 
                             test_prop = params$test_prop,
                             stratified_by = y)
Xtrain <- data_split$X$train
Xvalid <- data_split$X$validate
Xtest <- data_split$X$test
ytrain <- data_split$y$train
yvalid <- data_split$y$validate
ytest <- data_split$y$test
```

Provide summary statistics and/or figures of the three data sets to illustrate how similar (or different) they are.

```{asis, interactive_text = TRUE}

```

### Data Splitting Overview {.tabset .custom-tabs}

#### X Data Split {.unnumbered}

```{r X-data-split, add_new_line = TRUE}
plot_X_datasplit(Xtrain, Xvalid, Xtest)
```

#### Y Data Split {.unnumbered}

```{r y-data-split, add_new_line = TRUE}
plot_y_datasplit(ytrain, yvalid, ytest)
```

## Data Cleaning and Preprocessing

What steps were taken to clean the data? More importantly, why was the data cleaned in this way?

Discuss all inconsistencies, problems, oddities in the data (e.g., missing data, errors in data, outliers, etc.).

Record your preprocessing steps in a way such that if someone else were to reproduce your analysis, they could easily replicate and understand your steps.

It can be helpful to include relevant plots that explain/justify the choices that were made when cleaning the data.

If more than one preprocessing pipeline is reasonable, examine the impacts of these alternative preprocessing pipelines on the final data results.

Again, be as transparent as possible. This allows others to make their own educated decisions on how best to preprocess the data.

```{asis, interactive_text = TRUE}

```

```{r preprocess-data}
Xtrain <- log(Xtrain + 1) %>%
  removeConstantCols(verbose = 1) %>%
  removeDuplicateCols(verbose = 1) %>%
  filterByVar(max.p = 1000)
Xvalid <- log(Xvalid + 1)[, colnames(Xtrain)]
Xtest <- log(Xtest + 1)[, colnames(Xtrain)]
```


## Data Exploration

TODO: Add drag and drop feature in shiny version for other images

The main goal of this section is to give the reader a feel for what the data "looks like" at a basic level.

Provide plots that summarize the data and perhaps even plots that convey some smaller findings which ultimately motivate the main findings.

Provide additional plots representing remaining oddities after pre-processing if applicable.

Add summary statistics in accompanying tables (or in figures) for quick comparisons.

```{asis, interactive_text = TRUE}

```

### Training Data Overview {.tabset .custom-tabs}

```{r echo = FALSE}
dataOverview(Xtrain = Xtrain, ytrain = ytrain) %>%
  cat()
```

#### Summary Tables {.unnumbered}

```{r results="asis", add_new_line = TRUE}
# summary of types of features in training X
data_types(Xtrain = Xtrain, ytrain = ytrain)
```

```{r results="asis", add_new_line = TRUE}
# summary of features in training X, sorted by type
tab_ls <- data_summary(Xtrain = Xtrain, ytrain = ytrain, digits = 2, sigfig = F)
for (dt_name in names(tab_ls)) {
  simChef:::subchunkify(tab_ls[[dt_name]], i = chunk_idx, 
                        other_args = "results='asis'")
  chunk_idx <- chunk_idx + 1
}
```

#### X distribution {.unnumbered}

```{r, add_new_line = TRUE}
# plot X distribution
plot_X_distribution(Xtrain, "density")
```

#### Y distribution {.unnumbered}

```{r, add_new_line = TRUE}
# plot y distribution
plot_y_distribution(ytrain, "bar")
```

#### Feature Correlation {.unnumbered}

```{r, add_new_line = TRUE}
# correlation heatmap
keep_features <- sort(
  sample(1:ncol(Xtrain), min(ncol(Xtrain), 1000), replace = FALSE)
)
plotCorHeatmap(X = Xtrain[, keep_features], cor_type = "pearson", 
                      clust = TRUE, text_size = 0) +
  ggplot2::labs(x = "Features", y = "Features", fill = "Cor.") +
  simChef::pretty_ggplot_theme(x_text_angle = TRUE)
```

#### Feature Pair Plots {.unnumbered}

```{r, add_new_line = TRUE}
# pair plots
keep_features <- sort(
  sample(1:ncol(Xtrain), min(ncol(Xtrain), 6), replace = FALSE)
)
plotPairs(data = Xtrain, columns = keep_features, 
          color = ytrain, color_label = "y")
```

#### Marginal Associations {.unnumbered}

```{r, add_new_line = TRUE}
caret::featurePlot(x = Xtrain[, keep_features],
                   y = ytrain,
                   plot = if (is.factor(ytrain)) "box" else "scatter",
                   # strip = strip.custom(par.strip.text = list(cex = .7)),
                   scales = list(x = list(relation = "free"), 
                                 y = list(relation = "free")))
```

#### PCA {.unnumbered}

```{r, add_new_line = TRUE}
# dimension reduction plots
plotPCA(X = Xtrain, npcs = 3, color = ytrain, color_label = "y",
        center = T, scale = FALSE)$plot
```

# Prediction Modeling

TODO: add advice on which models to select and why

Discuss the prediction methods under consideration, and explain why these methods were chosen.

```{asis, interactive_text = TRUE}

```

Discuss the accuracy metrics under consideration, and explain why these metrics were chosen.

```{asis, interactive_text = TRUE}

```

Note: there should be multiple methods and metrics under consideration to paint a more holistic picture of the data. At least one method should be a baseline, common approach that may not be optimal for the problem setting, but serves as a helpful comparison.

## Prediction check

Carry out the prediction pipeline, outlined above.

-   Fit prediction methods on training data.
-   Evaluate prediction methods on validation data.
-   Compare results, and filter out poor models.

```{asis, interactive_text = TRUE}

```

```{r caret-fits, eval = params$modeling_pkg == "caret", echo = params$modeling_pkg == "caret"}
## Pick models and fitting backend for caret
tr_control <- caret::trainControl(
  method = "cv",
  number = 5,
  classProbs = if (is.factor(ytrain)) TRUE else FALSE,
  summaryFunction = caret::defaultSummary,
  allowParallel = FALSE,
  verboseIter = FALSE
)

response_type <- "raw"
model_list <- list(
  ranger = list(tuneGrid = expand.grid(mtry = seq(sqrt(ncol(Xtrain)), 
                                                  ncol(Xtrain) / 3,
                                                  length.out = 3),
                                       splitrule = "gini",
                                       min.node.size = 1),
                importance = "impurity",
                num.threads = 1),
  xgbTree = list(tuneGrid = expand.grid(nrounds = c(10, 25, 50, 100, 150),
                                        max_depth = c(3, 6),
                                        colsample_bytree = 0.33,
                                        eta = c(0.1, 0.3),
                                        gamma = 0,
                                        min_child_weight = 1,
                                        subsample = 0.6),
                 nthread = 1)
)

start_time <- Sys.time()
model_results <- fitCaret(Xtrain = Xtrain, ytrain = ytrain, 
                          Xtest = Xvalid, ytest = yvalid, 
                          model_list = model_list, tr_control = tr_control, 
                          response_type = response_type)
cat(sprintf("Fitting time taken: %f minutes", 
            difftime(Sys.time(), start_time, units = "mins")))
```

```{r tidymodels-fits, eval = params$modeling_pkg == "tidymodels", echo = params$modeling_pkg == "tidymodels"}
# TODO: add code for tuning parameters
# AK: Once tuning is fixed, can adjust splits for stability

## Pick models and fitting backend for tidymodels

# as is, only works for classification
rf_model <- parsnip::rand_forest() %>%
  parsnip::set_args(mtry = tune::tune()) %>%
  parsnip::set_engine("ranger", importance = "impurity") %>%
  parsnip::set_mode("classification")
rf_grid <- tidyr::crossing(mtry = 1:4)

svm_model <- parsnip::svm_rbf() %>%
  parsnip::set_engine("kernlab") %>%
  parsnip::set_mode("classification")

knn_model <- parsnip::nearest_neighbor() %>%
   parsnip::set_args(neighbors = tune(), weight_func = tune()) %>% 
   parsnip::set_engine("kknn") %>% 
   parsnip::set_mode("classification")

model_list <- list(RF = list(model = rf_model,
                             grid = rf_grid), 
                   SVM = list(model = svm_model,
                              grid = NULL), 
                   KNN = list(model = knn_model,
                              grid = 4))
kfolds <- 5

start_time <- Sys.time()
model_results <- fitTidyModels(Xtrain = Xtrain, ytrain = ytrain, 
                               Xtest = Xvalid, ytest = yvalid,
                               model_list = model_list, kfolds = kfolds)
cat(sprintf("Fitting time taken: %f minutes", 
            difftime(Sys.time(), start_time, units = "mins")))
```

```{r h2o-fits, eval = params$modeling_pkg == "h2o", echo = params$modeling_pkg == "h2o"}
## Pick models and fitting backend for h2o
model_list <- list(randomForest = list(ntrees = 500), 
                   xgboost = list())

start_time <- Sys.time()
model_results <- fith2o(Xtrain = Xtrain, ytrain = ytrain,
                        Xtest = Xvalid, ytest = yvalid,
                        model_list = model_list)
cat(sprintf("Fitting time taken: %f minutes", 
            difftime(Sys.time(), start_time, units = "mins")))
```

```{r prediction-accuracy-html, results = "asis", eval = knitr::is_html_output()}
# (html) table of accuracy metrics
model_results$errors %>%
  simChef::pretty_DT(digits = 2, sigfig = F, rownames = FALSE, 
           caption = "Validation Prediction Accuracies", 
           options = list(dom = 't'))
```

```{r prediction-accuracy-latex, results = "asis", eval = knitr::is_latex_output()}
# (latex) table of accuracy metrics
model_results$errors %>%
  simChef::pretty_kable(digits = 2, sigfig = F, row.names = FALSE, 
           caption = "Validation Prediction Accuracies",
           format = "latex")
```

## Stability check

Taking the prediction methods that pass the prediction check, perform stability analysis.

-   Specify and justify the appropriate data perturbation(s).
-   Re-fit the prediction methods on these perturbed data sets.
-   Evaluate prediction methods on validation data.
-   Assess stability across the data perturbations as well as across the various methods.
-   Filter out poor models where necessary and interpret stability results.

```{asis, interactive_text = TRUE}

```

TODO: Add results for h20, tuning needs to be fixed for h20 + tidymodels <!-- TODO: Ana - Provide some example code here (for both fitting and visualizing results) so that the practitioner can easily input their data and models. Add something like James' slide. A few pictures with possible data perturbation schemes (separate from parameter tuning). Data splitting vs sampling of observations. Cross-validation-ish scheme and fixed training/validation setup (bootstrapping, sub-sampling, stratified-sampling). Add parameter to include/exclude certain code chunks. -->

```{r stability-check}
n_reps <- 2 # increase for better stability measures when not testing code

# p <- progressr::progressor(steps = n_reps)
# future::plan(multisession, workers = min(n_reps, parallel::detectCores() - 1))
bootstrap_model_results <- future.apply::future_replicate(
  n = n_reps,
  expr = {
    bootstrap <- sample(1:nrow(Xtrain), nrow(Xtrain), replace = TRUE)
    Xtrain_b <- Xtrain[bootstrap, ]
    ytrain_b <- ytrain[bootstrap]
    
    if (params$modeling_pkg == "caret") {
      model_results_b <- fitCaret(Xtrain = Xtrain_b, ytrain = ytrain_b, 
                                  Xtest = Xvalid, ytest = yvalid, 
                                  model_list = model_list, 
                                  tr_control = tr_control, 
                                  response_type = response_type)
    } else if (params$modeling_pkg == "tidymodels") {
      model_results_b <- fitTidyModels(Xtrain = Xtrain_b, ytrain = ytrain_b, 
                                       Xtest = Xvalid, ytest = yvalid,
                                       model_list = model_list, kfolds = kfolds)
    } else if (params$modeling_pkg == "h2o") {
      model_results_b <- fith2o(Xtrain = Xtrain_b, ytrain = ytrain_b,
                                Xtest = Xvalid, ytest = yvalid,
                                model_list = model_list)
    }
    return(model_results_b)
  },
  simplify = FALSE
)

bootstrap_model_errs <- purrr::map_dfr(bootstrap_model_results, "errors", 
                                       .id = "bootstrap_id")
bootstrap_model_preds <- purrr::map(bootstrap_model_results, "predictions",
                                    .id = "bootstrap_id")
bootstrap_model_imps <- purrr::map(bootstrap_model_results, "importance",
                                   .id = "bootstrap_id") %>%
  dplyr::bind_rows(.id = "bootstrap_id")

bootstrap_model_errs_summary <- bootstrap_model_errs %>%
  dplyr::group_by(model) %>%
  dplyr::summarise(dplyr::across(-bootstrap_id, list(mean = mean, sd = sd))) %>%
  setNames(stringr::str_replace(colnames(.), "_", " "))
```

```{r stability-check-table-html, results = "asis", eval = knitr::is_html_output()}
# (html) table of accuracy metrics
bootstrap_model_errs_summary %>%
  simChef::pretty_DT(
    digits = 2, sigfig = F, rownames = FALSE, 
    caption = "Validation Prediction Accuracies Over Bootstrapped Training Fits", 
    bold_function = ". == max(., na.rm = TRUE)", bold_margin = 2,
    bold_scheme = c(F, T, F, T, F),
    options = list(dom = 't')
  )
```

```{r stability-check-table-latex, results = "asis", eval = knitr::is_latex_output()}
# (latex) table of accuracy metrics
bootstrap_model_errs_summary %>%
  simChef::pretty_kable(
    digits = 2, sigfig = F, row.names = FALSE, 
    caption = "Validation Prediction Accuracies Over Bootstrapped Training Fits", 
    bold_function = ". == max(., na.rm = TRUE)", bold_margin = 2,
    bold_scheme = c(F, T, F, T, F),
    format = "latex"
  )
```

```{r stability-check-boxplot}
# boxplots
bootstrap_model_errs %>%
  tidyr::pivot_longer(cols = -c(model, bootstrap_id), 
                      names_to = "Metric", values_to = "Value") %>%
  plotBoxplot(x_str = "Value", y_str = "model", horizontal = FALSE) +
  ggplot2::facet_wrap(~ Metric) +
  ggplot2::labs(x = "Method", 
                title = "Validation Prediction Accuracies Over Bootstrapped Training Fits")
```

## Interpretability {.tabset .custom-tabs}

For the models that pass the prediction and stability checks, extract the important features in the predictive models that are stable across both data and model perturbations. Determining the importance of a feature can be method dependent.

```{asis, interactive_text = TRUE}

```

### Full Model (without stability) {.unnumbered .tabset .tabset-pills}

#### Table {.unnumbered}

```{r add_new_line = TRUE, results = "asis", eval = knitr::is_html_output()}
simChef::pretty_DT(model_results$importance, digits = 2, sigfig = F, 
                   caption = "Variable Importances")
```

```{r add_new_line = TRUE, results = "asis", eval = knitr::is_latex_output()}
simChef::pretty_kable(model_results$importance, digits = 2, sigfig = F, 
                      caption = "Variable Importances", format = "latex")
```

#### Plots {.unnumbered}

```{r, add_new_line = TRUE}
plotFeatureImportance(model_results$importance,
                      use_rankings = FALSE,
                      use_facets = TRUE,
                      interactive = FALSE)
```

```{r, add_new_line = TRUE}
plotFeatureImportancePair(model_results$importance,
                          use_rankings = TRUE,
                          interactive = FALSE)
```

### Bootstrapped Model (with stability) {.unnumbered .tabset .tabset-pills}

#### Table {.unnumbered}

```{r add_new_line = TRUE, results = "asis", eval = knitr::is_html_output()}
bootstrap_model_imps_summary <- bootstrap_model_imps %>%
  dplyr::group_by(model, variable) %>%
  dplyr::summarise(`Mean Importance` = mean(importance),
                   `Median Importance` = median(importance),
                   `SD Importance` = sd(importance),
                   `Min Importance` = min(importance),
                   `Max Importance` = max(importance))
simChef::pretty_DT(bootstrap_model_imps_summary, 
                   digits = 2, sigfig = F, 
                   caption = "Summary of variable importances across bootstrapped models")
```

```{r add_new_line = TRUE, results = "asis", eval = knitr::is_latex_output()}
bootstrap_model_imps_summary <- bootstrap_model_imps %>%
  dplyr::group_by(model, variable) %>%
  dplyr::summarise(`Mean Importance` = mean(importance),
                   `Median Importance` = median(importance),
                   `SD Importance` = sd(importance),
                   `Min Importance` = min(importance),
                   `Max Importance` = max(importance))
simChef::pretty_kable(bootstrap_model_imps_summary, 
                      digits = 2, sigfig = F, 
                      caption = "Summary of variable importances across bootstrapped models",
                      format = "latex")
```

#### Plots {.unnumbered}

```{r, add_new_line = TRUE}
plotFeatureImportanceStability(bootstrap_model_imps,
                               use_rankings = FALSE,
                               use_facets = FALSE,
                               interactive = FALSE)
```

# Main Results

Interpret and summarize the prediction and stability results.

```{asis, interactive_text = TRUE}

```

Evaluate pipeline on test data.

```{r final-fits, eval = TRUE }
Xtrain_final <- dplyr::bind_rows(Xtrain, Xvalid)
ytrain_final <- c(ytrain, yvalid)
if (params$modeling_pkg == "caret") {
  final_model_results <- fitCaret(Xtrain = Xtrain_final, ytrain = ytrain_final, 
                                  Xtest = Xtest, ytest = ytest, 
                                  model_list = model_list, 
                                  tr_control = tr_control, 
                                  response_type = response_type)
} else if (params$modeling_pkg == "tidymodels") {
  final_model_results <- fitTidyModels(Xtrain = Xtrain_final, 
                                       ytrain = ytrain_final, 
                                       Xtest = Xtest, ytest = ytest,
                                       model_list = model_list, kfolds = kfolds)
} else if (params$modeling_pkg == "h2o") {
  final_model_results <- fith2o(Xtrain = Xtrain_final, ytrain = ytrain_final,
                                Xtest = Xtest, ytest = ytest,
                                model_list = model_list)
}

```

Summarize test set prediction and/or interpretability results.

```{asis, interactive_text = TRUE}

```

<!-- TODO: Ana - add template tables with interpretation -->

```{r test-accuracy-table-html, results = "asis", eval = knitr::is_html_output()}
# (html) table of accuracy metrics
final_model_results$errors %>%
  simChef::pretty_DT(digits = 2, sigfig = F, rownames = FALSE, 
           caption = "Test Prediction Accuracies", 
           options = list(dom = 't'))
```

```{r test-accuracy-table-latex, results = "asis", eval = knitr::is_latex_output()}
# (latex) table of accuracy metrics
final_model_results$errors %>%
    simChef::pretty_kable(digits = 2, sigfig = F, row.names = FALSE, 
             caption = "Test Prediction Accuracies", format = "latex")
```

# Post hoc analysis

Move beyond the global prediction accuracy metrics and dive deeper into individual-level predictions for the validation and/or test set, i.e., provide a more "local" analysis.

-   Examine any points that had poor predictions.
-   Examine differences between prediction methods.

```{asis, interactive_text = TRUE}

```

```{r}
plotPairs(cbind(`True Repsonses` = ytest, final_model_results$predictions), 
          columns = 1:(ncol(final_model_results$predictions) + 1), 
          title = "Comparison of model predictions")
```

# Conclusions

Reiterate main findings, note any caveats, and clearly translate findings/analysis back to the domain problem context.

```{asis, interactive_text = TRUE}

```

